Composing Robustly Safe Compilers

### 2023-06-09 popl/main.tex
Add overview

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
% To this date, no practical compiler implementation is proven as robustly safe.
% This is no surprise, since proving compilers robustly safe is a huge endeavor.
% But, compilers consist of several different passes, which themselves can be seen as compilers that translate from one intermediate representation to another.
%
% This paper investigates the compositionality of robustly safe compilers.
% Composing two compilers means to simply plug the output of one as an input to the other.
% We show that it suffices to prove individual passes as robustly safe.
% We demonstrate that one can simply prove robust memory safety preservation for one pass and prove the robust preservation of cryptographic constant time for another pass.
%
%  Robustly Safe Compilers ensure that programs that do not go wrong when linked with an arbitrary source-level attacker also do not go wrong in their compiled counterpart when linked with untrusted target-level code.
  Programs that do not go wrong should also not go wrong when linked with arbitrary, untrusted code.
  Robustly safe compilers ensure that, given such program at source level, this also holds at target level.
  The robust safety criterion requires challenging proofs which, so far, do not exist in real-world compiler implementations.
  Moreover, these implementations are modular:
  Compilers consist of several different so-called passes, which translate from one intermediate representation to another.
  It is an open question whether one can split the proof of robust safety modularily as well.

  This paper answers this question positively:
  The output of one robust safety preserving compiler may serve as input to another and this composition of both compilers remains robustly safe preserving.
  We generalise this compositional behaviour for compilers that robustly preserve different security properties.
  Thus, formal compiler engineers do not need to prove that all passes robustly preserve all security properties of interest, because their composition yields the intersection of the security properties.
  A consequence of this is that even in the context of the phase ordering problem, whenever the optimisations are proven robustly safe, we can freely swap the optimisation passes without compromising security.
  We validate our meta-level findings by means of a case study that looks at three different robust safety preserving compilers.

  Lastly, parts of our results are verified in Coq.
\end{abstract}

@@

\begin{lstlisting}[language=c,caption=``Wrong bounds check of two 64-bit integers.'']
if at <= bounds {
  x[at] = 42;
}
\end{lstlisting}
Note the subtle bug of using \verb|<=| instead of \verb|<|, leading to an out-of-bounds access whenever \verb|at = bounds|.
Furthermore, when \verb|at| and \verb|bounds| are 64-bit integers on a 32-bit architecture, the comparison may not be performed in constant-time: The compiler may bail out as soon as the lower 32 bits are unequal, not bothering to compare the higher 32 bits.

\section{Compositionality}

\section{Robust MS Preservation}

\section{Robust MS+CCT Preservation}

\section{Related Work}

\section{Conclusion}

%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
To Robert, for the bagels and explaining CMYK and color spaces.
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-base}

>>>>

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  \hltt{abstract}

\begin{center}\small\it
	{In the following, we use syntax highlighting accessible to both colourblind and black \& white readers~\citep{patrignani2020use}. Specifically, we use a $\src{\text{blue}}$,
	$\src{sans\text{-}serif}$ font for $\src{L_{TMS}}$,
	$\trg{bold}$ font for $\trg{L}$ and $\trg{L_{MS}}$, and $\obj{green}$ font for $\obj{L_{sCCT}}$.
	}
\end{center}
\end{abstract}

@@

\hltt{Context}
\hltt{Problem}

\BrText{Overview}{This paper introduces a formal framework for secure compiler compositions and showcases the framework's power by instantiating it on a multi-level compilation chain.
To this end, we discuss security property compositions and define several secure compilers, where each is either robustly preserving a different security property of interest or optimizing, i.e., applying \gls{cf} or doing \gls{dce}.
We compose these secure compilers into a multi-level compilation chain which is optimizing and robustly preserves a combination of \gls{ms} and \gls{cct}.
Crucially, the robust preservation proofs are only done for the individual compilers in the compilation chain and the robust preservation for the whole chain is a result of one of our key theorems.
The key results are formalized in Coq.}

In summary, this paper makes the following contributions:

\begin{itemize}
  \item
        \Cref{sec:compprop} presents a formal framework to reason about compositions of security properties and demonstrate its usage with concrete examples:
        Given a most general trace model, we define the trace-based security properties \gls{tms}, \gls{sms}, \gls{ms}, \gls{cct}, and their compositions.
        For each of these properties, we also give a monitor which runs on another trace model that excludes unneccessary information and show that their composition is behaviorally similar to the composition of the trace-based security properties.
        This two-level approach enhances modularity: When working with the security properties, one only needs to reason about the corresponding monitor.

  \item
        \Cref{sec:compcomp} studies different forms of secure compiler compositions.
        Concretely, it investigates upper and lower compositions, as well as the sequential composition.
        A simple corollary of sequential composition is the ability to freely reorder compiler optimizations without compromising security.

  \item \Cref{sec:casestud} sets up a case-study to demonstrate the power of our formal framework.
        We define several different languages and compilers between them and prove that the compilers are robustly preserving different properties.
        The sequential composition of these compilers yields a compilation chain that is optimizing and robustly preserving a combination of \gls{ms} and \gls{cct}.
        As for optimizations, the chain has two passes, one performing \gls{cf} and the other \gls{dce}.
        This compilation chain demonstrates the ability to perform modular proof engineering by leaveraging this paper's compositionality theorems.

  \item Our key contributions are formalized in the Coq proof assistant.
        The formalization for the compilers are \emph{extractable}.
        That is, one can translate our Coq code to Ocaml and run them.
        Finally, in this paper, whenever definitions or proofs are done in Coq, we use the symbol \CoqSymbol.
\end{itemize}

We now describe the other sections of the paper.
In \Cref{sec:background}, the paper establishes the notions of traces, (hyper-)properties, and robust preservation, while \Cref{sec:relwork} compares this work with others.
We conclude in \Cref{sec:concl} and give an outlook for future work.

\contribution{Open Source \& Technical Report} Our technical report and the Coq formalisation are
available as supplementary material and will be made available upon acceptance of this paper.

\section{Background}\label{sec:background}
We give a brief introduction to trace properties, monitors, and property satisfaction (\Cref{subsec:bg:tprop}).
With property satisfaction, we re-introduce the concept of robust satisfaction and its compiler-driven preservation (\Cref{subsec:bg:rtp}).

\subsection{Trace-Based Properties}\label{subsec:bg:tprop}
\subsection{Robust Trace-Property Preservation}\label{subsec:bg:rtp}

\section{Composing Properties}\label{sec:compprop}
\section{Composing Robustly Secure Compilers}\label{sec:compcomp}
This section presents three different kinds of secure compiler compositions and their corresponding proofs.

\section{Case Study: Memory Safe and Cryptographic Constant Time Code}\label{sec:casestud}
This section defines several secure compilers, each of which robustly preserves a different property of interest.
We demonstrate the power of our framework from \Cref{sec:compprop,sec:compcomp} by composing these compilers for a secure, but optimizing, compilation chain that robustly preserves \gls{ms} and \gls{cct}.
Concretely, \Cref{subsec:cs:tms} describes robust preservation of \gls{tms}. % by compiling from a language with a strong, static type system to a language without any type system.
\Cref{subsec:cs:ms} describes the robust preservation of \gls{sms} and, when composed with the compiler for \gls{tms}, \gls{ms}.
\Cref{subsec:cs:scct} composes the compiler for \gls{ms} with another one that robustly preserves \gls{cct}.
We attain our final compilation chain in \Cref{subsec:cs:opts}, which defines optimizing compilers performing either \gls{cf} or \gls{dce} and composes them with the ones from \Cref{subsec:cs:tms,subsec:cs:ms,subsec:cs:scct}.

\subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
\subsection{Robust (Spatial) Memory Safety Preservation}\label{subsec:cs:ms}
\subsection{Robust Strict Cryptographic Constant Time Preservation}\label{subsec:cs:scct}
\subsection{Optimizing Compilers}\label{subsec:cs:opts}

%\begin{lstlisting}[language=c,caption=``Wrong bounds check of two 64-bit integers.'']
%if at <= bounds {
%  x[at] = 42;
%}
%\end{lstlisting}
%Note the subtle bug of using \verb|<=| instead of \verb|<|, leading to an out-of-bounds access whenever \verb|at = bounds|.
%Furthermore, when \verb|at| and \verb|bounds| are 64-bit integers on a 32-bit architecture, the comparison may not be performed in constant-time: The compiler may bail out as soon as the lower 32 bits are unequal, not bothering to compare the higher 32 bits.

\section{Related Work}\label{sec:relwork}
\section{Conclusion}\label{sec:concl}

%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{main}
